Concentration of Maxima and Fundamental Limits in High-Dimensional Testing and Inference

### 2021-04-04 supplement/proofs-GWAS.tex
rename files and folders according to the new structure
many changes... Chapter 4 intro WIP

\section{Auxiliary facts of chi-square distributions}
\label{sec:chi-square-distributions}
\input{6.GWAS/6.chi-square-distributions.tex}

@@

The condition $\underline{r} > {{g}}(\beta)$ implies, after some algebraic manipulation,
$\sqrt{\underline{r}} -\sqrt{1-\beta} > 1$.
Therefore, we can pick $q>1$ such that 
\begin{equation} \label{eq:choice-of-q}
    \sqrt{\underline{r}} -\sqrt{1-\beta} > \sqrt{q} > 1.
\end{equation}
Setting the $t^* = t^*_p = 2q\log{p}$, we have $t_p < t^*_p$ for large $p$.

@@

The proof follows the ideas in \cite{arias2017distribution}, and is very similar to the proof of Theorem \ref{thm:Gaussian-error-approx-boundary}. 
One could in principle obtain the proofs in this section by referencing arguments that have appeared in Chapter \ref{chap:phase-transitions}.
We choose to present the proof here in full to make this section self-contained.

@@

As with the proof of Theorem \ref{thm:chi-squared-approx-boundary}, one could shorten the presentations in this section by 
referencing arguments in Chapter \ref{chap:phase-transitions}.  Again, we choose to present the proof in full to make this section 
self-contained.

@@

Since $\underline{r}>\widetilde{g}(\beta)=1$, we can pick $q$ such that $1<q<\underline{r}$.
Let $t^* = 2q\log{p}$, we have $t_p<t_p^*$ for large $p$ as in the proof of Theorem \ref{thm:chi-squared-exact-boundary}.
Therefore for large $p$, we have
$$
\widehat{W}_\text{signal}(t_p) \ge \widehat{W}_\text{signal}(t^*) \ge \overline{F_a}(t^*) + o_\P(1),
$$
where the last inequality follows from the stochastic monotonicity of the chi-square family (Lemma \ref{lemma:stochastic-monotonicity}), and Lemma \ref{lemma:empirical-process}.
Indeed, $F_a(t^*)\to0$ by \eqref{eq:approx-boundary-proof-sufficient-2} and our choice of $q<\underline{r}$. 
The proof of the sufficient condition is complete.

Proof of the necessary condition follows a similar structure to that of Theorem \ref{thm:chi-squared-approx-boundary}.
That is, we show that $\mathrm{FWER} + \mathrm{FNR}$ has liminf at least 1 by working with the lower bound
\begin{equation} \label{eq:exact-approx-boundary-proof-necessary-1}
    \mathrm{FWER}(\mathcal{R}) + \mathrm{FNR}(\mathcal{R}) \ge \P\left[\max_{i\in S^c}x(i)>u\right] \wedge \E\left[\frac{|S\setminus \widehat{S}(u)|}{|S|}\right],
\end{equation}
which holds for any thresholding procedure $\mathcal{R}$ and for arbitrary $u\in\R$.
By the assumption that $\overline{r}<\widetilde{g}(\beta)=1$, we can pick $q$ such that $\overline{r}<q<1$ and let $u = t^*=2q\log{p}$.
By relative stability of chi-squared random variables (Lemma \ref{lemma:rapid-variation-chisq}), we have
\begin{equation} \label{eq:exact-approx-boundary-proof-necessary-2}
    \P\left[\frac{\max_{i\in S^c} x(i)}{2\log{p}} > \frac{t^*}{2\log{p}}\right] \to 1.
\end{equation}
where the first fraction in \eqref{eq:exact-approx-boundary-proof-necessary-2} converges to 1, while the second converges to $q<1$.
On the other hand, by our choice of $q>\overline{r}$, the second term in \eqref{eq:exact-approx-boundary-proof-necessary-1} also converges to 1 as in \eqref{eq:approx-boundary-proof-converse-4}.
This completes the proof of the necessary condition.
\end{proof}

@@

By the assumption that $\underline{r}>\widetilde{h}(\beta)=(\sqrt{\beta}+\sqrt{1-\beta})^2$, we have $\sqrt{\underline{r}}-\sqrt{1-\beta}>\sqrt{\beta}$, so we can pick $q>0$, such that 
\begin{equation} \label{eq:approx-exact-boundary-proof-sufficient-2}
\sqrt{\underline{r}}-\sqrt{1-\beta}>\sqrt{q}>\sqrt{\beta}.
\end{equation}
Let $t^*=2q\log{p}$, we claim that 
\begin{equation} \label{eq:approx-exact-boundary-proof-sufficient-3}
\P\left[\tau\le t^*\right]\to 1.
\end{equation}
Indeed, by our choice of $q>\beta$, \eqref{eq:approx-exact-boundary-proof-sufficient-3} follows in the same way that \eqref{eq:approx-boundary-proof-sufficient-3} did.

@@

By the assumption that $\overline{r}<\widetilde{h}(\beta)=(\sqrt{\beta}+\sqrt{1-\beta})^2$, we can pick a constant $q>0$, such that 
\begin{equation} \label{eq:approx-exact-boundary-proof-necessary-2}
    \sqrt{\overline{r}} - \sqrt{1-\beta} < \sqrt{q} < \sqrt{\beta}.
\end{equation}
Let also $u = t^*=2q\log{p}$.
By our choice of $q < \beta$, we know from \eqref{eq:approx-boundary-proof-converse-3} that the first term on the right-hand-side of \eqref{eq:approx-exact-boundary-proof-necessary-1} converges to 1.
It remains to show that the second term in \eqref{eq:approx-exact-boundary-proof-necessary-1} also converges to 1.

>>>>

\section{Auxiliary facts of chi-square distributions}
\label{sec:chi-square-distributions}
\input{./supplement/chi-square-distributions.tex}

@@

The condition $\underline{r} > f_{\mathrm{E}}(\beta)$ implies, after some algebraic manipulation,
$\sqrt{\underline{r}} -\sqrt{1-\beta} > 1$.
Therefore, we can pick $q>1$ such that 
\begin{equation} \label{eq:choice-of-q}
    \sqrt{\underline{r}} -\sqrt{1-\beta} > \sqrt{q} > 1.
\end{equation}
Setting the $t^* = t^*_p = 2q\log{p}$, we have $t_p < t^*_p$ for large $p$.

@@

The proof follows the ideas in \cite{arias2017distribution}, and is very similar to the proof of Theorem \ref{thm:Gaussian-error-approx-boundary}. 
One could in principle obtain the proofs in this section by referencing arguments that have appeared in Chapter \ref{chap:phase-transitions}.
We choose to present the proof here in full for completeness.

@@

As with the proof of Theorem \ref{thm:chi-squared-approx-boundary}, one could shorten the presentations in this section by referencing arguments in Chapter \ref{chap:phase-transitions}.  
%Again, we choose to present the proof in full to make this section self-contained.

@@

Since $\underline{r}>f_{\mathrm{EA}}(\beta)=1$, we can pick $q$ such that $1<q<\underline{r}$.
Let $t^* = 2q\log{p}$, we have $t_p<t_p^*$ for large $p$ as in the proof of Theorem \ref{thm:chi-squared-exact-boundary}.
Therefore for large $p$, we have
$$
\widehat{W}_\text{signal}(t_p) \ge \widehat{W}_\text{signal}(t^*) \ge \overline{F_a}(t^*) + o_\P(1),
$$
where the last inequality follows from the stochastic monotonicity of the chi-square family (Lemma \ref{lemma:stochastic-monotonicity}), and Lemma \ref{lemma:empirical-process}.
Indeed, $F_a(t^*)\to0$ by \eqref{eq:approx-boundary-proof-sufficient-2} and our choice of $q<\underline{r}$. 
The proof of the sufficient condition is complete.

Proof of the necessary condition follows a similar structure to that of Theorem \ref{thm:chi-squared-approx-boundary}.
That is, we show that $\mathrm{FWER} + \mathrm{FNR}$ has liminf at least 1 by working with the lower bound
\begin{equation} \label{eq:exact-approx-boundary-proof-necessary-1}
    \mathrm{FWER}(\mathcal{R}) + \mathrm{FNR}(\mathcal{R}) \ge \P\left[\max_{i\in S^c}x(i)>u\right] \wedge \E\left[\frac{|S\setminus \widehat{S}(u)|}{|S|}\right],
\end{equation}
which holds for any thresholding procedure $\mathcal{R}$ and for arbitrary $u\in\R$.
By the assumption that $\overline{r}<f_{\mathrm{EA}}(\beta)=1$, we can pick $q$ such that $\overline{r}<q<1$ and let $u = t^*=2q\log{p}$.
By relative stability of chi-squared random variables (Lemma \ref{lemma:rapid-variation-chisq}), we have
\begin{equation} \label{eq:exact-approx-boundary-proof-necessary-2}
    \P\left[\frac{\max_{i\in S^c} x(i)}{2\log{p}} > \frac{t^*}{2\log{p}}\right] \to 1.
\end{equation}
where the first fraction in \eqref{eq:exact-approx-boundary-proof-necessary-2} converges to 1, while the second converges to $q<1$.
On the other hand, by our choice of $q>\overline{r}$, the second term in \eqref{eq:exact-approx-boundary-proof-necessary-1} also converges to 1 as in \eqref{eq:approx-boundary-proof-converse-4}.
This completes the proof of the necessary condition.
\end{proof}

@@

By the assumption that $\underline{r}>f_{\mathrm{AE}}(\beta)=(\sqrt{\beta}+\sqrt{1-\beta})^2$, we have $\sqrt{\underline{r}}-\sqrt{1-\beta}>\sqrt{\beta}$, so we can pick $q>0$, such that 
\begin{equation} \label{eq:approx-exact-boundary-proof-sufficient-2}
\sqrt{\underline{r}}-\sqrt{1-\beta}>\sqrt{q}>\sqrt{\beta}.
\end{equation}
Let $t^*=2q\log{p}$, we claim that 
\begin{equation} \label{eq:approx-exact-boundary-proof-sufficient-3}
\P\left[\tau\le t^*\right]\to 1.
\end{equation}
Indeed, by our choice of $q>\beta$, \eqref{eq:approx-exact-boundary-proof-sufficient-3} follows in the same way that \eqref{eq:approx-boundary-proof-sufficient-3} did.

@@

By the assumption that $\overline{r}<f_{\mathrm{AE}}(\beta)=(\sqrt{\beta}+\sqrt{1-\beta})^2$, we can pick a constant $q>0$, such that 
\begin{equation} \label{eq:approx-exact-boundary-proof-necessary-2}
    \sqrt{\overline{r}} - \sqrt{1-\beta} < \sqrt{q} < \sqrt{\beta}.
\end{equation}
Let also $u = t^*=2q\log{p}$.
By our choice of $q < \beta$, we know from \eqref{eq:approx-boundary-proof-converse-3} that the first term on the right-hand-side of \eqref{eq:approx-exact-boundary-proof-necessary-1} converges to 1.
It remains to show that the second term in \eqref{eq:approx-exact-boundary-proof-necessary-1} also converges to 1.
